---
title: "Network annotations"
author: "C.deVriend"
date: "23-6-2022"
output: html_document
toc: true
toc_float: true
number_sections: true
theme: readable
code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Network annotations
At this point I have created a general network from the CARNIVAL analysis by unionizing the nodes and edges from each subset. In order to allow for better interpretation the network will be copied 3 times, where I will project the mean expression for set 1 (Day 0-1.5, associated with the self-renewal state), set 2 (Day 1.5-3, assoicated with intermedeate differentiation) and set 3 (Day 3-6, associated with terminal differentiation). 

```{r}
# Load the weighted sif into excel
Combined_weighted_sif_useful.xlsx <- read.csv("C:/Users/cheye/OneDrive/Bureaublad/Radboud/Master 21-22/Internship/Data and analysis/CARNIVAL/analysis/Combined_weighted_sif_useful.xlsx.csv", sep=";")
# Remove "column 1"/first column
Combined_weighted_sif_useful.xlsx <-  Combined_weighted_sif_useful.xlsx[ -c(1) ]

# Load VST normalized counts 
VST_normalized_counts <- read_excel("VST_normalized_counts.xlsx")
# column to rownames
VST_normalized_counts<-column_to_rownames(VST_normalized_counts, var = "rowname")
# Convert to matrix
#VST_normalized_counts<-as.matrix(VST_normalized_counts)
# transform
#VST_normalized_counts<-t(VST_normalized_counts)
# scale the columns
#VST_normalized_counts <- scale(VST_normalized_counts)
# transform back
#VST_normalized_counts<-t(VST_normalized_counts)
# back to df
#VST_normalized_counts<-as.data.frame(VST_normalized_counts)



# Create new df that has the VST normalized MEAN counts per set (these are set 1, 2 and 3 I've explained earlier)
Normalized_Expression_Set1_2_3<-VST_normalized_counts
Normalized_Expression_Set1_2_3$Set1<-rowMeans(subset(VST_normalized_counts, select = c(d00_b1,d00_b2, d05_b1, d05_b2, d10_b1,d10_b2,d15_b1,d15_b2)), na.rm = TRUE)
Normalized_Expression_Set1_2_3$Set2<-rowMeans(subset(VST_normalized_counts, select = c(d15_b1,d15_b2, d20_b1, d20_b2, d25_b1,d25_b2,d30_b1,d30_b2)), na.rm = TRUE)
Normalized_Expression_Set1_2_3$Set3<-rowMeans(subset(VST_normalized_counts, select = c(d30_b1,d30_b2, d45_b1, d45_b2, d50_b1,d50_b2,d60_b1,d60_b2)), na.rm = TRUE)
# Give rownames
#Normalized_Expression_Set1_2_3<-column_to_rownames(Normalized_Expression_Set1_2_3, var = "rowname")
# Remove columns we don't need
Normalized_Expression_Set1_2_3<- Normalized_Expression_Set1_2_3[ -c(1:20) ]
# Scale columns
Normalized_Expression_Set1_2_3<- scale(t(Normalized_Expression_Set1_2_3))
# Transform back
Normalized_Expression_Set1_2_3<-t(Normalized_Expression_Set1_2_3)


# Create string of node 1 and node 2 so we can extract the names from Normalized expression set1/2/3
Node_names_total_network<- Combined_weighted_sif_useful.xlsx[,1,3] 
# Select only unique names
#Node_names_total_network<-unique(Node_names_total_network)

# We have 3 interacting complexes in the network so lets create a row in the expression matrix by creating the mean expression between the interacting genes per set1/2/3
# First we find the rownumbers for the genes of interest
 which(rownames(Normalized_Expression_Set1_2_3) == "CCNE2") # ROW 11982
  which(rownames(Normalized_Expression_Set1_2_3) == "CDK2") # 4829
   which(rownames(Normalized_Expression_Set1_2_3) == "SEC61A1") #716
    which(rownames(Normalized_Expression_Set1_2_3) == "SEC61A2")#866
     which(rownames(Normalized_Expression_Set1_2_3) == "SEC61B") #3048
          which(rownames(Normalized_Expression_Set1_2_3) == "SEC61G") #5862
      which(rownames(Normalized_Expression_Set1_2_3) == "SMAD3") #10518
       which(rownames(Normalized_Expression_Set1_2_3) == "SMAD4") #7256

# quickly create unique dfs for each set so life becomes easier
set1<-as.data.frame(Normalized_Expression_Set1_2_3)  
set1<-set1[ -c(2:3) ]
set2<-as.data.frame(Normalized_Expression_Set1_2_3)
set2<-set2[ -c(1,3) ]
set3<-as.data.frame(Normalized_Expression_Set1_2_3)
set3<-set3[-c(1:2)]

# SEC61A1/2/1B/1G
mean(set1[c(716,866,3048,5826),])
mean(set2[c(716,866,3048,5826),])
mean(set3[c(716,866,3048,5826),])
# 0.4696541, 0.3475734, -0.8172276

# as.data.frame
Normalized_Expression_Set1_2_3<-as.data.frame(Normalized_Expression_Set1_2_3)
# Add new row to existing df
Normalized_Expression_Set1_2_3[nrow(Normalized_Expression_Set1_2_3) + 1,] <- c(0.4696541, 0.3475734, -0.8172276)
# Change the rowname
rownames(Normalized_Expression_Set1_2_3)[rownames(Normalized_Expression_Set1_2_3) == "22457"] <- "SEC61A1_SEC61A2_SEC61B_SEC61G"

# SMAD3/4 
mean(set1[c(10518,7256),])
mean(set2[c(10518,7256),])
mean(set3[c(10518,7256),])
# 0.390923, -0.5566802,0.1657572

# Add new row to existing df
Normalized_Expression_Set1_2_3[nrow(Normalized_Expression_Set1_2_3) + 1,] <- c(0.390923, -0.5566802,0.1657572)
# Change the rowname
rownames(Normalized_Expression_Set1_2_3)[rownames(Normalized_Expression_Set1_2_3) == "22458"] <- "SMAD3_SMAD4"

# CCNE2/CDK2
mean(set1[c(11982,4829),])
mean(set2[c(11982,4829),])
mean(set3[c(11982,4829),])
# 1.107653, -0.8242975, -0.2833552

# Add new row to existing df
Normalized_Expression_Set1_2_3[nrow(Normalized_Expression_Set1_2_3) + 1,] <- c(1.107653, -0.8242975, -0.2833552)
# Change the rowname
rownames(Normalized_Expression_Set1_2_3)[rownames(Normalized_Expression_Set1_2_3) == "22459"] <- "CCNE2_CDK2"

# Finally create the new df that only contains the extracted names from the node name string
Node_expressions<-Normalized_Expression_Set1_2_3[rownames(Normalized_Expression_Set1_2_3) %in% c(Node_names_total_network), ] # 10 node expressions not present!!

# Write excel file from node expressions to use it in cytoscape
Node_expressions<-rownames_to_column(Node_expressions, var = "Node_names")
writexl::write_xlsx(Node_expressions, "Node_Expressions_total_union.xlsx")

# It appeared that some nodes were missed extracting it this way so I will do it manually # 161 nodes of # 174 (13 nodes expressions missing)
Node_expressions<-Normalized_Expression_Set1_2_3[rownames(Normalized_Expression_Set1_2_3) %in% c("ADNP",
"AHR",
"AIP",
"AR",
"ARID3A",
"ARNT",
"ARNTL",
"ASCL1",
"ATE1",
"ATF1",
"ATF2",
"ATF3",
"ATF4",
"ATF6",
"ATF7",
"BACH1",
"BAG4",
"BCL6",
"BHLHE40",
"CADPS",
"CADPS2",
"CAMK1",
"CCNE2_CDK2",
"CDK1",
"CDK2",
"CDK20",
"CDK2AP1",
"CDK5",
"CDK5RAP3",
"CDKN3",
"CDX2",
"CEBPA",
"CEBPB",
"CEBPD",
"CEBPE",
"CEBPG",
"CERK",
"CHEK1",
"CHP1",
"CLOCK",
"CPLX1",
"CREB1",
"CREB3",
"CREB3L1",
"CREB5",
"CREBZF",
"CREM",
"CTCF",
"CUX1",
"DNAJB9",
"DTX1",
"DUSP10",
"DUSP14",
"DUSP19",
"DUSP2",
"DUSP26",
"DUSP7",
"DUSP8",
"DUSP9",
"E2F2",
"E2F3",
"E2F4",
"E2F5",
"E2F6",
"EBF1",
"EFNA1",
"EFNA2",
"EFNA3",
"EFNA4",
"EFNA5",
"EFNB2",
"EFNB3",
"EGFR",
"EGR1",
"EID2",
"ELF1",
"ELF3",
"ELK1",
"ELK4",
"ELP1",
"EPAS1",
"EPHA4",
"ERBB2",
"ESR1",
"FBXL12",
"FBXO11",
"FBXO22",
"FBXW11",
"FHL5",
"FOXH1",
"GABPB1",
"GABPB2",
"GXYLT1",
"GXYLT2",
"HCFC1",
"HDLBP",
"HEY2",
"HIF1AN",
"HNF1B",
"HSPA5",
"HYAL2",
"IGBP1",
"ISG15",
"ITCH",
"JAK1",
"JAK2",
"KDM7A",
"KRT14",
"LAMTOR3",
"LCMT1",
"LTA",
"MAP3K6",
"MAPK1",
"MAPK14",
"MAPK3",
"MAPK8",
"MBTPS2",
"MDK",
"MTMR4",
"N4BP1",
"NAB1",
"NAB2",
"NOTCH2",
"NR0B2",
"NRG1",
"NUMBL",
"NUP214",
"OMG",
"PAK1",
"PCBD1",
"PCSK7",
"PDHX",
"POGLUT1",
"PPM1F",
"PPP2CA",
"PPP2CB",
"PPP2R5C",
"PRKACA",
"PRKCH",
"PRPF4B",
"PTP4A1",
"PTP4A2",
"PTPRB",
"PTPRE",
"PTPRG",
"RAF1",
"RB1",
"RBCK1",
"RBL2",
"RORB",
"SEC61A1_SEC61A2_SEC61B_SEC61G",
"SIM2",
"SMAD2",
"SMAD3_SMAD4",
"SMAD4",
"SNIP1",
"SOCS4",
"STK40",
"STX1A",
"TAF12",
"TAF4",
"TAOK2",
"TFDP2",
"TGIF1",
"TIPRL",
"TLX1",
"TNFRSF1A",
"TNFRSF1B",
"TP53",
"TRIM33",
"TUB",
"USP9X",
"WDR83",
"ZNF521"),]

# Write excel file from node expressions to use it in cytoscape
Node_expressions<-rownames_to_column(Node_expressions, var = "Node_names")
writexl::write_xlsx(Node_expressions, "Node_Expressions_total_union_2.xlsx")
```

Another approach is to extract the nodes unique to the three subset union views I created in cytoscape. This will be the approach I will be following from now on and is probably going to be included into my final report.
```{r}
# Extract nodes from unionized subsets 1-3
Nodes_subset_union_1.3 <- read.csv("C:/Users/cheye/OneDrive/Bureaublad/Radboud/Master 21-22/Internship/Data and analysis/CARNIVAL/data/Union subset 1-3/Nodes_subset_union_1-3.csv")
# Create vector containing node names
Node_names_union_1.3<- Nodes_subset_union_1.3$name
# Extract node expressions from scaled vst data
Node_Expressions_Unionized_Subsets_1_2_3_nodesset1<-Normalized_Expression_Set1_2_3[rownames(Normalized_Expression_Set1_2_3) %in% c(Node_names_union_1.3),] 
# ALL 85 node expression were sucesfully extracted; let's write an excel file
Node_Expressions_Unionized_Subsets_1_2_3_nodesset1<-rownames_to_column(Node_Expressions_Unionized_Subsets_1_2_3_nodesset1, var = "name")
writexl::write_xlsx(Node_Expressions_Unionized_Subsets_1_2_3_nodesset1, "Node_Expressions_Unionized_Subsets_1_2_3_nodesset1.xlsx")

# Extract nodes from unionized subsets 4-6
Nodes_subset_union_4.6 <- read.csv("C:/Users/cheye/OneDrive/Bureaublad/Radboud/Master 21-22/Internship/Data and analysis/CARNIVAL/data/Union subset 1-3/Nodes_subset_union_4-6.csv")
# Create vector containing node names
Node_names_union_4.6<- Nodes_subset_union_4.6$name
# Extract node expressions from scaled vst data
Node_Expressions_Unionized_Subsets_1_2_3_nodesset2<-Normalized_Expression_Set1_2_3[rownames(Normalized_Expression_Set1_2_3) %in% c(Node_names_union_4.6),] 
# 82 genes from 88 extracted; let's write an excel file
Node_Expressions_Unionized_Subsets_1_2_3_nodesset2<-rownames_to_column(Node_Expressions_Unionized_Subsets_1_2_3_nodesset2, var = "name")
writexl::write_xlsx(Node_Expressions_Unionized_Subsets_1_2_3_nodesset2, "Node_Expressions_Unionized_Subsets_1_2_3_nodesset2.xlsx")

# Extract nodes from unionized subsets 7-9
Nodes_subset_union_7.9 <- read.csv("C:/Users/cheye/OneDrive/Bureaublad/Radboud/Master 21-22/Internship/Data and analysis/CARNIVAL/data/Union subset 1-3/Nodes_subset_union_7-9.csv")
# Create vector containing node names
Node_names_union_7.9<- Nodes_subset_union_7.9$name
# Extract node expressions from scaled vst data
Node_Expressions_Unionized_Subsets_1_2_3_nodesset3<-Normalized_Expression_Set1_2_3[rownames(Normalized_Expression_Set1_2_3) %in% c(Node_names_union_7.9),] 
# 75 genes from 82 extracted; let's write an excel file
Node_Expressions_Unionized_Subsets_1_2_3_nodesset3<-rownames_to_column(Node_Expressions_Unionized_Subsets_1_2_3_nodesset3, var = "name")
writexl::write_xlsx(Node_Expressions_Unionized_Subsets_1_2_3_nodesset3, "Node_Expressions_Unionized_Subsets_1_2_3_nodesset3.xlsx")


```










